# Reproducible Research Week 2 Course Project


### Loading data
Data loaded using read.csv()
```{r}
data<-read.csv("activity.csv")
head(data)
library(knitr)
library(dplyr)
library(ggplot2)
```


### What is mean total number of steps taken per day?

1. Summing steps of each day into variable totalstepsperday
```{r}
StepsPerDay<-aggregate(data$steps,list(data$date),sum)
colnames(StepsPerDay)<-c("Date","Steps")
```

2. Creating a histogram of Date vs Total Steps
```{r}
hsteps<-ggplot(StepsPerDay,aes(Steps))
hsteps+geom_histogram(col="brown",fill="orange")+ggtitle("Steps Per Day HISTOGRAM")+xlab("Steps")+ylab("Frequency Count")
```

3. Mean of the number of steps taken per day
```{r}
smean<-mean(StepsPerDay$Steps,na.rm = TRUE)
smean
```

4. Median of the number of steps taken per day
```{r}
smedian<-median(StepsPerDay$Steps,na.rm = TRUE)
smedian
```


###What is the average daily activity pattern?

1. Making time series plot of steps vs intervals
```{r}
StepsPerInterval<-aggregate(steps~interval,data,mean,na.action=na.omit)
tsplot<-ggplot(StepsPerInterval,aes(interval,steps))
tsplot+geom_line(col="darkgreen")+ggtitle("Steps vs Intervals")+xlab("Intervals")+ylab("Steps")
```

2. The 5-minute interval having maximum no. of steps
```{r}
max<-filter(StepsPerInterval,steps==max(StepsPerInterval$steps))
```


###Imputing missing values
1. Calculating and reporting total no. of NA values 
```{r}
mvals<-sum(is.na(data))
mvals
```

2. Replacing missing values 
```{r}
data$nsteps<-ifelse(is.na(data$steps),round(StepsPerInterval$steps[match(data$interval,StepsPerInterval$interval)],0),data$steps)
```

3. New dataset with filled in missing values
```{r}
NewData<-data.frame(steps=data$nsteps,date=data$date,interval=data$interval)
head(NewData)
```

4. Histogram of steps per day with the new dataset 
```{r}
NewStepsPerDay<-aggregate(NewData$steps,list(NewData$date),sum)
colnames(NewStepsPerDay)<-c("NewDate","NewSteps")

Newhsteps<-ggplot(NewStepsPerDay,aes(NewSteps))
Newhsteps+geom_histogram(color="blue",fill="cyan")+ggtitle("Histogram with new dataset")+xlab("New Steps")+ylab("Frequency Count")

NewMean<-mean(NewStepsPerDay$NewSteps)
NewMean

NewMedian<-median(NewStepsPerDay$NewSteps)
NewMedian
```
Comparing both the histograms, we come to know that between 10000 and 12500 steps, there is a sudden rise in frequency because of the replacement of missing values. With a slight variation, the mean and median of the data have decrease since for some values of steps, the frequencies have increased.

###Are there differences in activity patterns between weekdays and weekends?

1.Creating new factor variable to check the day type 
```{r}
NewData$ActualDate<-as.Date(NewData$date,format="%Y-%m-%d")
NewData$WeekDay<-weekdays(NewData$ActualDate)
NewData$TypeOfDay<-ifelse(NewData$WeekDay=='Saturday' | NewData$WeekDay=='Sunday', 'weekend','weekday')
head(NewData)
```

2. Panel Plot for average steps taken with respect to 5-minute intervals on weekday and weekend 
```{r}
NewStepsPerInterval<-aggregate(steps~interval+TypeOfDay,NewData,mean,na.action=na.omit)
Newtsplot<-ggplot(NewStepsPerInterval,aes(interval,steps))
Newtsplot+geom_line(col="red")+ggtitle("Steps vs Intervals")+xlab("Intervals")+ylab("Steps")+facet_grid(TypeOfDay~.)
```